\documentclass[twoside,12pt]{mythesis} %this is the mythesis.cls file
% twoside,openright,versioninfo

% Consider:
% \newcommand{\ie}{i.\,e.}
% \newcommand{\Ie}{I.\,e.}
% \newcommand{\eg}{e.\,g.}
% \newcommand{\Eg}{E.\,g.}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{pdfpages} % new addition
\usepackage{xparse}
\usepackage{float}
\usepackage{setspace} % new addition

\usepackage{geometry} % new addition 
 \geometry{
 a4paper,
 total={210mm,297mm},
 left=35mm,
 right=20mm,
 top=20mm,
 bottom=20mm,
 }

% http://www.tug.dk/FontCatalogue/allfonts.html
\usepackage[T1]{fontenc}
\usepackage[default]{gillius}

%\usepackage[utf8]{inputenc}
%\usepackage{tgadventor} %% Added to change font
 


% \usepackage{xltxtra} % new addition
%\usepackage{cite}
% \usepackage{epigraph} can add an inspirational quote with this

 %\usepackage[round]{natbib}

%You might need to load other packages here...

% \RequirePackage{refstyle}
% \renewcommand{\figref}{\Figref}
% \renewcommand{\tabref}{\Tabref}

%% Symbols
% unresolved:
\newcommand{\ud}{\mathrm{d}}
\newcommand{\drel}{\ensuremath{r_{\mathrm{rel}}}}
\newcommand{\nmax}{\ensuremath{n_\mathrm{max}}}

%Information for the title page
%Some of this is hard coded in mythesis.cls but you can over write if you need to
\title{The \uppercase{e}cology of \uppercase{o}bligate \uppercase{s}cavengers from \uppercase{i}ndividual \uppercase{b}ehaviour to \uppercase{p}opulation \uppercase{d}ynamics}
%
\author{Adam Kane}
%
\month{\textsc{January}} \year{2015}
\previousdegrees{B.Sc.(Hons), University College Dublin, 2007\\%
  M.Sc., Dublin City University, 2010}
\degreetitle{Doctor of Philosophy}
\institution{Trinity College Dublin}
\school{School of Natural Sciences}
\department{Zoology}

%End of preamble. 

\begin{document}
%\raggedright

%-------------------------------------
% \fontfamily{qag}\selectfont

% added this to change font
%-------------------------------------
\maketitle %puts in your title

\chapter*{Declaration}
I declare that this thesis has not been submitted as an exercise for a degree at this or any other university and it is, unless otherwise referenced, entirely my own work.
I agree to deposit this thesis in the University's open access institutional repository or allow the library to do so on my behalf, subject to Irish Copyright Legislation and Trinity College Library conditions of use and acknowledgement.

\vspace{10 mm}

\noindent{Adam Kane}

\chapter*{Summary}
Studies on vultures are on the rise, and just as well given the sharp declines in many of the 23 species. Indeed it seems these population crashes are responsible for this research boost. However, there remain obvious gaps in our knowledge when it comes to the world's only terrestrial example of vertebrate obligate scavengers. It has been suggested that a human aversion to carrion is one of the reasons scavengers are reviled by the public and understudied by science. Broadly, the following body of work is an attempt to fill in some of these gaps. To achieve this I use a variety of methods and look at a range of species. I start by giving a description of the impressive ranging abilities of one species, the Cape Vulture, in Southern Africa, and identify causes of variation in these home ranges. I then move to Kenya where I identify a producer-scrounger game occuring between the \textit{Gyps} vultures and scavenging eagles of the area. I use a suite of empirical data and modelling approaches to show vultures use eagles in identifying the location of carcasses. The results of which underscore the importance of ecosystem-level management. Then in chapter 3, I collect data on the food required by the avian scavenging guild of Swaziland and use a novel method, namely population dynamics P systems, to ascertain if the wild fauna of the country can meet this need. Although sufficient for the time being carrion will soon become limiting and I advocate the creation of supplementary feeding sites in the country to counter this. Finally, I visit the Cretaceous Period where I use my understanding of modern day carrion ecology to explore a putative scavenger in \textit{Tyrannosaurus rex}, the infamous carnivorous dinosaur. A proposed hypothesis that \textit{T.rex} took on more carrion in its diet across ontogeny is shown to be false owing to the huge cost of movement as the animal grew. 

\chapter*{Acknowledgements}
The infamous 'they' warn that, with a bad supervisor, a PhD is a hellish experience. Thank God then for Andrew Jackson who made the previous three years of 'work' one of the most enjoyable periods of my life. The wonderful atmosphere of the zoology department certainly helped too. Luke McNally and Mafalda Viana guided me from my first year as the senior PhDs in the office. Deirdre McClean and Se{\'a}n Kelly were my early morning catalysts whose mania ensured no Monday morning was ever a drudge. While Kevin Healy's cynicism and creativity complemented this well. The academic odd couple of Thomas Guillerme and Sive Finaly along with their supervisor Natalie Cooper were always there to help me get unstuck. Joe Colgan deserves mention for advertising the department in such an appealing light; I was encouraged to start a PhD given his experience (Alain Finn take note). All the other members of NERD club, I thank you for stimualting discussions. \\
\indent
In the field Ara Monadjem is owed a debt for preventing mine on more than one occasion which made our collaborations a lot more successful. Mduduzi Ngwenya and the members of All Out Africa all helped to navigate me through my time in Swaziland. \\
\indent
Thanks too to my co-authors outside of Trinity College, Kerri, Walter, Graeme, Antoni, M.Angels, and Darcy. You provided the data and helped me with the analysis, all of which made this piece of work possible. \\
\indent
My mam and dad were responsible for getting me interested in science in the first place and for it they, along with my brothers Alex and Tom, have suffered/listened to someone who has no problem telling them how interesting it is. \\
\indent
Dank je wel to Gwendoline, we began at the same time as the PhD, and she has made my time outside of zoology just as enjoyable as my time spent within. 


%\input{abstract} %inputs abstract.tex
%\input{preface} %inputs preface.tex
\allcontents %tells it to make a table of contents with figure and table lists too
%\input{acknowledgements} %inputs acknowledgements.tex
\cleardoublepage
\mainbody

\input{general_intro} %This is a ch1-introduction.tex file with contents of intro chapter

\input{ch1-movement} 

\input{ch2-scrounger}
	 
\input{ch3-pdp}

\input{ch4-trex}

\input{discussion}
 %\pagestyle{empty}% attempt to get rid of chapter 6 headers in bib
\formatbibliography
\bibliographystyle{PLoS-Biology} 
\bibliography{bibfile} 


%\formatbibliography %note there are only a few refs in this example in the appendix
%\bibliographystyle{refstyle} %you may want to make a new reference style. This is refstyle.bst
%\bibliography{refs-thesis} %you may also call you bibliography something different. This is refs-thesis.bib



\formatappendices
\input{ch2-scrounge-supp}
\input{ch3-pdp-supp}
\input{other-papers}


\end{document}
